---
title: "Viz I"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
library(hexbin)
```

Import the weather data

```{r}
data("weather_df")

```

Making our first plot :-)

```{r}

ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()

# need to specify which geom I want to use
# + used for plots instead of piping

```

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

```

```{r}

ggp_weather_scatterplot = 
  weather_df |>
    ggplot(aes(x = tmin, y = tmax)) +
    geom_point()

ggp_weather_scatterplot

```

Check that some rows are missing.

```{r}

weather_df |>
  filter(is.na(tmax))

# produce summary of how much missingness is in the data
# naniar can also be used to explore and manage missing values

```

## Fancier scatterplots!

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) # creates lines

```

Where you define aesthetics can matter

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE) # only one smooth is included

```

use faceting real quick

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) # shows all three groups (names) as separate plots, column version

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .) # row version

```

Let's make a somewhat more interesting scatterplot

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

# different shapes are given to the names


```

Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).

LA plot:

```{r}

weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax * (9/5) + 32,
    tmin_fahr = tmin * (9/5) + 32
  ) |>
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
    geom_point() +
    geom_smooth(se = FALSE, method = "lm")


```

## Small things

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
#  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)
# you don't need to have points in the plot

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_smooth(se = FALSE) +
  geom_point() 
# here the points come after the smooth


```

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()

```

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
    geom_point(color = "#086bfd")
# can choose the color of the points


```

## Univariate plots

```{r}

weather_df |>
  ggplot(aes(x = tmin)) +
  geom_histogram(color = "white", fill = "red")
# color aesthetic is the founds, fill is the color of the bins

```

```{r}

weather_df |>
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram()

weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram()

```

how would i fix this (better see the distribution of the data)? maybe facet?

```{r}

weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(name ~ .) # not overlapping anymore

```

maybe a density plot?

```{r}

weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.2) # made the color lighter to see the overlap better


```

```{r}

weather_df |>
  ggplot(aes(x = name, y = tmin)) +
    geom_boxplot(aes(fill = name))


```

violin plots

```{r}

weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
    geom_violin()
# violin plot is a vertical density plot that is symmetrical
# good to tell you about the modality


```

ridge plot

```{r}

weather_df |>
  ggplot(aes(x = tmin, y = name, fill = name)) +
  geom_density_ridges()
# good for showing density

```

LA univariate plots

Learning Assessment: Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.

```{r}

weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.2)

weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram() +
  facet_grid(name  ~ .)

weather_df |>
  filter(prcp > 5, prcp < 1000) |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.2)
# zooming in on the data

```

## Saving and embedding plots

saving plots

```{r}

ggp_weather_violin = 
  weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
    geom_violin()

ggp_weather_violin

ggsave("plots/violin_plot.pdf", ggp_weather_violin,
       width = 8, height = 6)
# ggsave() takes plot name first, then plot you want to save, can also choose width and height

```

embedding plots

```{r, fig.width = 6, fig.height = 4}

ggp_weather_violin

```


